---
title: Random commitment model qualitative comparison
date: "`r Sys.Date()`"
author: "fredcallaway"
output:
  rmdformats::robobook:
    code_folding: hide
    self_contained: true
  output:
    pdf_document:
      latex_engine: xelatex
---

```{r setup, include=FALSE}
source("setup.r")
source("load_data.r")
raw_df = bind_rows(
    read_sim("optimal_prior", noise_sd=3),
    read_sim("optimal_prior_high", noise_sd=3),
    read_sim("optimal", noise_sd=3),
)

df = raw_df %>% 
    filter(
        response_type %in% c("correct", "timeout") &
        abs(rel_strength) < 4.24 &
        abs(strength_first) < 3 &
        abs(strength_second) < 3
    ) %>%
    mutate(
        # name = recode_factor(name, .ordered=T,
        #     # "Optimal" = "optimal",
        #     "optimal_prior" = "Optimal",
        #     "human" = "Human",
        #     "empirical" = "Random",
        #     "empirical_commitment" = "Random Commitment",
        #     # "rand_gamma" = "Random",
        # ),
        last_pres = if_else(n_pres %% 2 == 1, "first", "second")
    )

```

```{r, fig.width=7.5, fig.height=3, include=FALSE}
make_fixations = function(df) {
    df %>% 
        ungroup() %>% 
        filter(n_pres >= 1) %>% 
        select(name, rel_strength, presentation_times, n_pres) %>% 
        mutate(
            trial = row_number(),
            strength_diff = cut(abs(rel_strength), 
                                c(0, 0.35, 1.25, 10),
                                # quantile(abs(rel_strength), c(0, 0.2, 1.),  na.rm = T),
                                labels=c("small", "moderate", "large"),
                                ordered=T)
        ) %>% 
        unnest_longer(presentation_times, "duration", indices_to="presentation") %>% 
        mutate(
            last_fix = as.numeric(presentation == n_pres),
            fix_first = presentation %% 2,
            fix_stronger = as.numeric(fix_first == (rel_strength > 0)),
            # duration = if_else(name == "Human", as.integer(duration), as.integer(duration * 250)),
        )
}

long = df %>% 
    group_by(name) %>%
    slice_sample(n=10000) %>% 
    make_fixations
```

# Main Results

Note: unless otherwise stated, all regression tables are maximal linear mixed effects models
(random slopes and intercepts) applied to the human data. 
Plotted regression lines are fixed effects only.

## Attentional time course

The key model prediction is that the probability of looking at the cue
with greater memory strength increases over time.

```{r, time-course, fig.width=7.5, fig.height=3.5}

normalized_timestep = function(long) {
    long %>% 
        group_by(trial) %>%
        mutate(prop_duration = duration / sum(duration)) %>% 
        ungroup() %>% 
        mutate(n_step=round(prop_duration * 100)) %>% 
        uncount(n_step) %>% 
        group_by(trial) %>% 
        mutate(normalized_timestep = row_number())
}

nts = normalized_timestep(long)
nts %>% 
    drop_na(strength_diff) %>% 
    ggplot(aes(normalized_timestep/100, fix_stronger, group = strength_diff, color=strength_diff)) +
    geom_smooth(se=F) + 
    ylim(0, 1) +
    facet_grid(~name) +
    labs(x="Normalized Time", y="Probability Fixate Stronger Cue", color="Strength Difference") +
    geom_hline(yintercept=0.5) +
    theme(legend.position="top")
```

In both the optimal model simulations and the human data, the probability of fixating
the stronger cue steadily increases over the course of the trial, and this increase
is more pronounced when there is a larger difference in strength.

In the random model, we only see a slight spike at the end of the trial. This
spike is due to the fact that the the model always remembers the item it looks
at last, and it is also more likely to remember the item with a stronger
memory.

The random commitment model does not show this effect, because the item it
remembers (and fixates on last) is primarly driven by the random decision to
commit to one of the items.

## Individual fixation durations

Although the random models we designed did not demonstrate the smoothly
increasing probability shown by both humans and the optimal model, it is
possible that the aforementioned last-fixation phenomenon could be partially
driving that result. We can avoid this confound by looking at the duration of
individual fixations, conditioning on the fact that each is not the final
fixation. The model predicts that fixations should be longer for stronger
cues, and (for non-initial fixations) when the competing cue is weaker.

In all these plots and analyses, final fixations are excluded.

### First fixation

```{r, first-fix, fig.width=7.5, fig.height=3}
df %>% 
    filter(n_pres >= 2) %>% 
    ggplot(aes(strength_first, first_pres_time)) +
    # geom_smooth(method="lm") + 
    geom_smooth() + 
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs(x="First Cue Memory Strength", y="First Fixation Time")

human %>% 
    filter(n_pres >= 2) %>% 
    lmer(first_pres_time ~ strength_first + (strength_first|wid), data=.) %>% summ
```

The optimal model predicts a weak but robust effect, while humans do not show any effect.

*NEW*: Previously, the model predicted a strong effect here.
The difference is primarly due to changing the distribution from which $p$
is drawn. When the distribution over $p$ is uniform (as before), each sample provides a lot
of information about which cue is stronger, and thus the model adapts very 
quickly. By using a prior that is skewed in one direction

By using a sharper prior, each sample provides less information,
and thus the model is not as sensitive to memory strength on the first fixation.


### Second fixation

```{r, second-fix, fig.width=7.5, fig.height=3}
df %>% 
    filter(n_pres >= 3) %>% 
    ggplot(aes(rel_strength, second_pres_time)) +
    geom_smooth(method="lm") + 
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs(x="Relative Memory Strength", y="Second Fixation Time")

X = human %>% 
     filter(n_pres >= 3) 

X %>% lmer(second_pres_time ~ rel_strength + (rel_strength|wid), data=.) %>% summ
# X %>% lmer(second_pres_time ~ strength_first + strength_second + (strength_first + strength_second|wid), data=.) %>% summ
```

The model makes a strong prediction here, and I don't think it is confounded
by anything. Previously, we had a significant effect here in the human data.
But excluding the incorrect trials pushed the p up a bit. It does seem
plausible that the effect is there though.

### Third fixation

```{r, third-fix, fig.width=7.5, fig.height=3}
random %>% filter(
    response_type == "correct",
    choose_first != (last_pres == "first")
)

df %>% 
    filter(n_pres >= 4) %>% 
    ggplot(aes(rel_strength, third_pres_time)) +
    geom_smooth(method="lm") + 
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs(x="Relative Memory Strength", y="Third Fixation Time")

human %>% 
    filter(n_pres >= 4) %>% 
    lmer(third_pres_time ~ rel_strength + (rel_strength|wid), data=.) %>% summ
```

I previously thought this could be due to selection effects, but given that
we aren't seeing anything in the random model, I think this result might be solid.



## Fixation proportion by relative memory strength

The simplest test of rational memory: do people spend more time looking at the cue
that they have a stronger memory of? This analysis only considers trials where
both cues are seen.

```{r, fix-prop, fig.width=7.5, fig.height=3}
df %>% 
    filter(n_pres >= 2) %>% 
    ggplot(aes(rel_strength, prop_first)) + 
    geom_smooth(method = "lm", formula="y~x") +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs(x="Relative Memory Strength", y="Proportion Fixate First")

X = df %>% filter(n_pres >= 2 & name == "Human")
lmer(choose_first ~ rel_strength + (rel_strength|wid), data=X) %>% summ
```

```{r, fix-prop, fig.width=7.5, fig.height=3}
df %>% 
    filter(n_pres >= 2) %>% 
    ggplot(aes(rel_strength, as.numeric(last_pres == "first"))) + 
    geom_smooth(method = "lm", formula="y~x") +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name)

# X = df %>% filter(n_pres >= 2 & name == "Human")
# lmer(choose_first ~ rel_strength + (rel_strength|wid), data=X) %>% summ
```

They do! 🎉 But hold on, there's a small but significant effect in the random model.
What's going on there?

### Last fixation effect

People remember the thing they look at last
`r round(100 * mean(multi$choose_last_seen, na.rm=T))`%
of the time. In value-based decision making, it has been suggested that
this last-fixation effect could explain away what looks to be evidence for
adaptive attention allocation. Briefly, the last fixation is
correlated with both fixation proportion and relative strength, and this
could be entirely driving the correlation between strength and fixation proportion.
See [here](last-fixation-21-06-01.html) for a 
more thorough explanation.

```{r, last-fix, fig.width=7.5, fig.height=3.5}
df %>% 
    filter(n_pres >= 2) %>% 
    ggplot(aes(rel_strength, prop_first, color=last_pres)) + 
    geom_smooth(method="lm") + 
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    theme(legend.position="top") +
    labs(x="Relative Memory Strength", y="Proportion Fixate First Cue", color="Last Fixation")

X = df %>% filter(n_pres > 1 & name == "Human")
lmer(prop_first ~ rel_strength * last_pres + (rel_strength * last_pres | wid), data=X) %>% summ

```

The effect of relative memory strength on fixation proportion is a lot weaker when
we control for the effect of the last fixation, but it remains significant. However,
it's _not_ significant in the optimal model (regression not shown).


## Final vs non-final duration

Longer final fixations suggest that people commit to a cue at some point. This
is a cool finding because we see the opposite (shorter final fixations) in
value-based and perceptual decisions, typically explained as the result of
crossing a threshold cutting off the final fixations. The reason we see this
in the optimal model is that the accumulation is not just to identify
which target is more memorable and also to actually remembering it (perceptual
and value-based decisions only have the former).

```{r, fix-by-final, fig.width=7.5, fig.height=3}
long %>% 
    # filter(name == "Human") %>% 
    ggplot(aes(last_fix==1, duration)) + 
    stat_summary(fun.data=mean_cl_boot, geom="bar", fill="white", color="black") +
    stat_summary(fun.data=mean_cl_boot, geom="errorbar", width=0.2) +
    facet_grid(~name) + 
    scale_x_discrete(name="Fixation Type", labels=c("Non-final", "Final"))
```

The random model result is interesting. The intention of the random
commmitment model was to capture this exact phenomenon. But it seems to have
way *over*-captured it. The reason this happens is (probably) that humans
don't actually randomly commit to a cue, but instead only commit when they are
reasonably sure they will remember it. Also, any such commitment is probably
not really a firm commitment. Anyway, what happens to this random model is
that it commits too early or on a bad cue, and then it takes forever to
remember it.

This might draw into question my decision to sample the commitment decision
based on the empirical number-of-fixations distribution. The only alternative
I can think of is to sample from a parametric distribution with some free 
parameter. Previously we discussed a probability of commiting on each fixation
which is equivalent to an Exponential distribution. But we don't have a good 
way to fit parameters...


# Sanity checks for evidence accumulation

All these plots are testing really basic predictions of the evidence
accumulation model, that is, effects that come out in the random model. They
are all much weaker in humans than the model and some aren't even significant.
This suggests that we might need to make some adjustments to the base
accumulation model, especially if we wanted to do any real model fitting.

## Probability of remembering first word

```{r,remember-first}
min(df$rel_strength, na.rm=T)
max(df$rel_strength, na.rm=T)
df %>%
    filter(response_type == "correct") %>% 
    ggplot(aes(rel_strength, as.numeric(choose_first))) + 
    geom_smooth(method = "glm", method.args = list(family = "binomial"), formula=y~x) +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    labs(x="Relative Memory Strength", y="Prob Select First Cue") +
    facet_wrap(~name)

human %>% 
    lmer(choose_first ~ rel_strength + (rel_strength|wid), data=.) %>% summ
# lmer(choose_first ~ strength_first + strength_second + (strength_first + strength_second |wid), data=X) %>% summ
```

## Reaction time by chosen strength

```{r, fig.width=7.5, fig.height=3}
df %>% 
    ggplot(aes(chosen_strength, rt)) +
    geom_smooth() +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name)

lmer(rt ~ chosen_strength + (chosen_strength|wid), data=human) %>% summ
```

## Last fixation duration by strength

```{r, fig.width=7.5, fig.height=3}
X = df %>% 
    filter(n_pres > 0) %>% 
    mutate(
        last_pres_time = map_dbl(presentation_times, last),
        last_rel_strength = if_else(last_pres == "first", rel_strength, 1 - rel_strength),
        last_strength = if_else(last_pres == "first", strength_first, strength_second)
    )

X %>% 
    ggplot(aes(last_strength, last_pres_time,)) +
    geom_smooth() +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs()

X %>% filter(name == "Human") %>% 
    lmer(last_pres_time ~ last_strength + (last_strength|wid), data=.) %>% summ
```

## Last fixation duration by strength and presentation

```{r, fig.width=7.5, fig.height=3}
X = df %>% 
    filter(between(n_pres, 1, 4)) %>% 
    mutate(
        last_pres_time = map_dbl(presentation_times, last),
        last_rel_strength = if_else(last_pres == "first", rel_strength, 1 - rel_strength),
        last_strength = if_else(last_pres == "first", strength_first, strength_second)
    )

X %>% 
    ggplot(aes(last_strength, last_pres_time, group=n_pres, color=n_pres)) +
    geom_smooth() +
    stat_summary_bin(fun.data=mean_cl_boot, bins=5) + 
    facet_grid(~name) +
    labs()

X %>% filter(name == "Human") %>% 
    lmer(last_pres_time ~ last_strength * n_pres + (1|wid), data=.) %>% summ
```



# Random stuff

## Timeout rate

```{r, fig.width=7.5, fig.height=3}
df %>% 
    ggplot(aes(response_type == "timeout", ..prop..)) +
    geom_bar(group=0) +
    facet_grid(~name)

```

## Reaction time

```{r, fig.width=7.5, fig.height=3}
df %>% 
    ggplot(aes(rt)) +
    geom_density() +
    facet_grid(~name)
```
## Number of fixations

```{r, fig.width=7.5, fig.height=3}
df %>% 
    filter(n_pres < 10) %>% 
    ggplot(aes(n_pres, ..prop..)) +
    geom_bar() +
    facet_grid(~name)

```

## Final fixation - relative strength

```{r, fig.width=7.5, fig.height=3}
X = df %>% 
    filter(n_pres > 3) %>% 
    mutate(
        last_pres_time = map_dbl(presentation_times, last),
        last_rel_strength = if_else(last_pres == "first", rel_strength, 1 - rel_strength),
        last_strength = if_else(last_pres == "first", strength_first, strength_second)
    )

X %>% ggplot(aes(last_strength, last_pres_time)) + 
    stat_summary_bin(fun.data=mean_cl_boot, bins=10) +
    facet_grid(~name) +
    geom_smooth(method='lm')

X %>% lmer(last_pres_time ~ last_rel_strength + (last_rel_strength|wid), data=.) %>% summ
```




