

# Sanity checks for evidence accumulation

All these plots are testing really basic predictions of the evidence
accumulation model, that is, effects that come out in the random model. 
They are all significant in the human data, but the RT ones are a lot
weaker. This suggests that we might need to make some adjustments to the base
accumulation model, especially if we wanted to do any real model fitting.

All these plots exclude timeout trials.

## Probability of remembering first word

```{r,remember-first, fig.width=WIDTH, fig.height=HEIGHT}
df %>%
    filter(response_type == "correct") %>% 
    mutate(choose_first = as.numeric(choose_first)) %>% 
    regress(rel_strength, choose_first) +
    ylab("Prob Select First Cue")
```

## Reaction time by chosen strength

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(response_type == "correct") %>% 
    regress(chosen_strength, rt)
```

## Last fixation duration by strength

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(response_type == "correct") %>% 
    filter(n_pres > 0) %>% 
    mutate(
        last_pres_time = map_dbl(presentation_times, last),
        last_rel_strength = if_else(last_pres == "first", rel_strength, 1 - rel_strength),
        last_strength = if_else(last_pres == "first", strength_first, strength_second)
    ) %>% 
    regress(last_strength, last_pres_time)
```

# Miscellaneous

## Timeout rate

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    group_by(name) %>% 
    summarise(timeout_rate = mean(response_type == "timeout")) %>% 
    kable(digits=2)
```

## Reaction time

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    ggplot(aes(rt)) +
    geom_density() +
    facet_grid(~name) +
    labs(x="Reaction Time", y="Density")
```

## Number of fixations

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres < 10) %>% 
    ggplot(aes(n_pres, ..prop..)) +
    geom_bar() +
    facet_grid(~name) +
    labs(x="Number of Fixations", y="Proportion of Trials")

```
