---
title: Rational metacognition in memory
date: "`r Sys.Date()`"
author: "fredcallaway"
output:
  rmdformats::robobook:
    code_folding: hide
    self_contained: true
---

```{r setup, include=FALSE}
DROP_HALF = FALSE
DROP_ACC = TRUE
DROP_ERROR = FALSE
source("setup.r")
source("load_data.r")
```

```{r}
if (DROP_ACC) cat(glue('Dropping {N_drop_acc} participants with less than 50% accuracy in critical trials'))
if (DROP_HALF) cat("Dropping the first half of critical trials")
if (DROP_ERROR) cat("Dropping error trials") else cat("_Including_ error trials")
```

Experiment versions: `r paste(VERSIONS)`

- Two rounds of training and pre-test
- No feedback in any pre-test trials
- Word stimuli are selected randomly (not selecting high/low recall probability and concreteness)
- Distractor task comes before final pre-test
- Strength is measured in the last round only
<!-- - **_Not_ excluding incorrect trials** -->


# Fixation time course

```{r, time-course, fig.width=WIDTH, fig.height=3}
long = df %>% 
    group_by(name) %>%
    slice_sample(n=10000) %>% 
    make_fixations

normalized_timestep = function(long) {
    long %>% 
        group_by(trial) %>%
        mutate(prop_duration = duration / sum(duration)) %>% 
        ungroup() %>% 
        mutate(n_step=round(prop_duration * 100)) %>% 
        uncount(n_step) %>% 
        group_by(trial) %>% 
        mutate(normalized_timestep = row_number())
}

long %>% 
    normalized_timestep %>% 
    drop_na(strength_diff) %>% 
    ggplot(aes(normalized_timestep/100, fix_stronger, group = strength_diff, color=strength_diff)) +
    geom_smooth(se=F) + 
    ylim(0, 1) +
    facet_grid(~name) +
    labs(x="Normalized Time", y="Probability Fixate\nStronger Cue", color="Strength Difference") +
    geom_hline(yintercept=0.5) +
    theme(legend.position="top")
```

# Fixation durations

## First fixation

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres >= 2) %>% 
    regress(strength_first, first_pres_time) # plot
```

## Second fixation

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres >= 3) %>% 
    regress(rel_strength, second_pres_time, bins=9) # plot
```

## Third fixation

```{r, third-fix, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres >= 4) %>% 
    regress(rel_strength, third_pres_time)
```

# Overall fixation proportion

```{r, fix-prop, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres >= 2) %>% 
    regress(rel_strength, prop_first)
```

## Last-fixation effect

```{r, last-confound, fig.width=WIDTH, fig.height=3}
df %>% 
    filter(n_pres >= 2) %>% 
    regress_interaction(rel_strength, last_pres, prop_first)
```

## Strength predicts last fixation 

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    mutate(last_pres_first = as.numeric(last_pres == "first")) %>% 
    regress(rel_strength, last_pres_first)
```

## Last fixations are longer

```{r, fix-by-final, fig.width=WIDTH, fig.height=HEIGHT}
long %>% 
    # filter(name == "Human") %>% 
    ggplot(aes(last_fix==1, duration)) + 
    stat_summary(fun.data=mean_cl_boot, geom="bar", fill="white", color="black") +
    stat_summary(fun.data=mean_cl_boot, geom="errorbar", width=0.2) +
    facet_grid(~name) + 
    scale_x_discrete(name="Fixation Type", labels=c("Non-final", "Final")) +
    ylab("Duration")

last_diff = long %>% 
    filter(name == "Human") %>% 
    with(tapply(duration, last_fix, mean)) %>% 
    diff

long %>% 
    filter(name == "Human") %>% 
    mutate(is_last = int(last_fix==1)) %>% 
    lmer(duration ~ is_last + (is_last|wid), data=.) %>% 
    summ

```


# Sanity checks for evidence accumulation

These plots test basic predictions of the evidence
accumulation model, that is, effects that come out in the random model. 

These plots only include correct trials.

## Probability of remembering first word

```{r,remember-first, fig.width=WIDTH, fig.height=HEIGHT}

df %>%
    filter(response_type == "correct") %>% 
    mutate(choose_first = int(choose_first)) %>% 
    regress(rel_strength, choose_first) +
    ylab("Prob Select First Cue")
```

## Reaction time by chosen strength

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(response_type == "correct") %>% 
    regress(chosen_strength, rt)
```

## Last fixation duration by strength

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(response_type == "correct") %>% 
    filter(n_pres > 0) %>% 
    mutate(
        last_pres_time = map_dbl(presentation_times, last),
        last_rel_strength = if_else(last_pres == "first", rel_strength, 1 - rel_strength),
        last_strength = if_else(last_pres == "first", strength_first, strength_second)
    ) %>% 
    regress(last_strength, last_pres_time)
```

# Miscellaneous

## Response types

Note: The model can only predict correct answers and timeouts.

```{r}
raw_df %>% 
    with((proportions(table(name, response_type), margin=1))) %>% 
    kable(digits=2)

```

## Reaction time

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    ggplot(aes(rt)) +
    geom_density() +
    facet_grid(~name) +
    labs(x="Reaction Time", y="Density")
```


## Number of fixations

```{r, fig.width=WIDTH, fig.height=HEIGHT}
df %>% 
    filter(n_pres < 10) %>% 
    ggplot(aes(n_pres, ..prop..)) +
    geom_bar() +
    facet_grid(~name) +
    labs(x="Number of Fixations", y="Proportion of Trials")

```

