
# Other junk

Below is a bunch of other stuff we tried that isn't really worth looking at.

## Timeouts

```{r}
all_trials %>% summarise(mean(response_type == "timeout"))
```

## AFC accuracy

```{r}

ggplot(trials, aes(afc_accuracy, log_recall_rt)) + 
    stat_summary()

trials %>% with(lm(log_recall_rt ~ afc_accuracy)) %>% summ

```
This looks promising, but it turns out this one is driven entirely by individual differences (each line is a participant).
Mixed effects regression shows no effect of 2AFC trials, especially when we account for RT on the correct 2AFC trials.

```{r}
trials %>%
    group_by(wid, afc_accuracy) %>% 
    summarise(log_recall_rt=mean(log_recall_rt)) %>% 
    ggplot(aes(afc_accuracy, log_recall_rt, group=wid)) + 
        geom_line() + theme(legend.position = "none")

trials %>% with(lmer(log_recall_rt ~ afc_accuracy + (1|wid))) %>% summ
trials %>% with(lmer(log_recall_rt ~ afc_accuracy + log_afc_rt + (1|wid))) %>% summ
```

## Time course

```{r}
ggplot(afc, aes(round, log_afc_rt)) + 
    stat_summary()

ggplot(afc, aes(round, as.numeric(correct))) + 
    stat_summary()
```

2AFC RT goes down, as we would expect.

```{r}
df = afc %>%
    select(wid, word, round, log_afc_rt) %>%
    pivot_wider(names_from=round, values_from=log_afc_rt, names_prefix='log_afc_rt') %>% 
    inner_join(select(trials, wid, word, log_recall_rt))

# lmer(log_recall_rt ~  log_afc_rt1 + log_afc_rt2 + (1|wid), data=df) %>% summ
```

The slope appears to be constant over rounds.

```{r}
select(df, starts_with('log')) %>% cor(use="pairwise.complete.obs") %>% kable
```

But the correlation is higher in rounds 2 and 3. (Look at the last column in the correlation matrix)



# Word type effects

## Accuracy

### All participants (before exclusion)
```{r}
all_trials %>% group_by(wid) %>% summarise(accuracy=mean(correct)) %>%
    ggplot(aes(accuracy)) + 
        geom_histogram(binwidth=.1)
```

```{r}
ggplot(all_trials, aes(fill=response_type, x=word_type)) +
     geom_histogram(stat="count") + response_type_colors
```


### Good participants (after exclusion)

```{r}
ggplot(trials, aes(fill=response_type, x=word_type)) +
     geom_histogram(stat="count") + response_type_colors
```

```{r, echo=TRUE}
glmer(correct ~ word_type + (1|wid), data=trials, family=binomial) %>% summ
```


## Reaction time

```{r}
trials %>% filter(rt < 5000) %>% ggplot(aes(rt)) + geom_histogram()
```

### Aggregate

```{r}
ggplot(trials, aes(x=response_type, y=rt, color=word_type)) + 
    stat_summary(fun.data=mean_se, geom="pointrange", position = position_dodge(width = 0.1)) + 
    theme(legend.position=c(0.2, 0.9)) + word_type_colors
trials %>% group_by(response_type, word_type) %>% summarise(rt = mean(rt)) %>% kable
```

### By participant

```{r}
grouped = trials %>% 
    group_by(wid, word_type, response_type) %>% 
    summarise(across(where(is.numeric), mean)) %>% 
    ungroup()

ggplot(grouped, aes(x=word_type, y=rt, color=response_type, group=response_type)) + 
    geom_line(alpha=0.3, size=1, aes(group=interaction(response_type, wid))) +
    stat_summary(fun.data=mean_se, geom="pointrange") + 
    stat_summary(fun.data=mean_se, geom="line", size=1) + 
    facet_wrap(~response_type) +
    response_type_colors + theme(legend.position="none")

ggplot(filter(grouped, response_type == "correct"), aes(x=word_type, y=rt, color=response_type, group=response_type)) + 
    geom_line(alpha=0.3, size=1, aes(group=interaction(response_type, wid))) +
    stat_summary(fun.data=mean_se, geom="pointrange") + 
    stat_summary(fun.data=mean_se, geom="line", size=1) + 
    response_type_colors + theme(legend.position="none")
```

```{r}
prop = grouped %>%
    filter(response_type == "correct") %>%
    group_by(wid, word_type) %>%
    summarise(rt=mean(rt)) %>% 
    pivot_wider(names_from=word_type, values_from=rt) %>% 
    with(round(mean(Low > High) * 100))
## prop
```

`r prop`% of participants respond faster on high memorability trials (including only correct response trials).


### Stats

```{r, echo=TRUE}
lmer(rt ~ word_type + (1|wid), data=trials) %>% summ
```

```{r, echo=TRUE}
lmer(rt ~ word_type * response_type + (1|wid), data=trials) %>% summ
```

```{r, echo=TRUE}
lmer(rt ~ word_type + (1|wid), data=subset(trials, correct)) %>% summ
```

```{r, echo=TRUE}
lmer(log_recall_rt ~ word_type + (1|wid), data=subset(trials, correct)) %>% summ
```

## Typing time

```{r}
ggplot(trials, aes(x=response_type, y=type_time, color=word_type)) + 
    stat_summary(fun.data=mean_se, geom="pointrange", position = position_dodge(width = 0.1)) + 
    theme(legend.position=c(0.2, 0.9)) + word_type_colors
trials %>% group_by(response_type, word_type) %>% summarise(type_time = mean(type_time)) %>% kable
```

```{r, echo=TRUE}
lm(type_time ~ word_type, data=subset(trials, correct)) %>% summ
```

## Total response time

```{r}
ggplot(trials, aes(x=response_type, y=rt+type_time, color=word_type)) + 
    stat_summary(fun.data=mean_se, geom="pointrange", position = position_dodge(width = 0.1)) + 
    theme(legend.position=c(0.2, 0.9)) + word_type_colors
trials %>% group_by(response_type, word_type) %>% summarise(total_time = mean(rt+type_time)) %>% kable
```

```{r, echo=TRUE}
lm(rt + type_time ~ word_type, data=subset(trials, correct)) %>% summ
``` 